SSA based global code motion of identical computations

### 2017-04-14 hoist.tex
merge cost-model and profitability

\subsection{Profitability check}
\label{subsec:profitability}
After the legality checks have passed, we check for profitability of hoisting.
That takes into account the impact code-hoisting would have on various
parameters that affect runtime performance e.g., impact on live-range, gain in
the code size.  We have established a set of cost models described in
Section~\ref{sec:cost-models} for each parameter and tuned them for performance
against representative benchmarks.

@@

\subsection{Hoisting bottom up with pessimistic approach}
\label{subsec:pessimistic}
In the pessimistic approach, the basic blocks are traversed in the inverse
depth-first order, computing the \GVN{} of instructions as they come by. The \GVN{} of
sibling branches are compared for equality. Once such a candidate is found, it
is hoisted in the common dominator. All the leaf nodes are visited before the
non-leaf nodes (bottom-up) because instructions are hoisted upwards.

The pessimistic algorithm is fast, results in fewer spills and also hoists fewer
instructions. We have implemented optimistic approach because it is more general
and can be tuned down to closely mimic pessimistic approach by changing a flag.

@@

\begin{verbatim}
B0: b = m
    goto B1

B1: c = n
    goto B2

B2: if c is true then goto B3 else goto B4

B3: a0 = b<kill> + c<kill>

B4: a1 = b<kill> + c<kill>

After hoisting a0 and a1 are removed and a copy
of a0 as a01 is placed in B2 just before the branch.
\end{verbatim}

\subsection{Hoisting too far away}
If there are several instructions in between the hoisting point and the
instruction to be hoisted, the instruction to be hoisted crosses several
instructions while hoisting, that means we are adding one register to all the
live-ranges spanning the instructions. That could result in spills. In the
current implementation we choose to hoist if the number of instructions crossed
is below a threshold. Ideally, it should be okay to hoist all the instructions
and a later a live-range-splitting \cite{cooper1998live} pass should make the
right decision of rematerializing the instruction should it be beneficial to do
so. But the current live-range splitting pass of \LLVM{} is not making the
optimial decision and we have found spills if the threshold is exceeded. The
threshold was computed as a result of tuning the \LLVMTestSuite{}, SPEC Cpu2000,
and Cpu2006 benchmarks \cite{Henning2000}.

Also, hoisting a load increases the register pressure by one across all the
instructions which the load would cross. That could result in spills later in
the register allocation.

However, in the special case discussed in Section~\ref{hoist:reg-pressure}, it
will be okay to hoist because the register pressure would not increase.

In order to reduce register pressure, we have restricted hoisting of address
computations away from their corresponding loads and stores when the loads and
stores cannot be moved.  This restriction is lifted when aggressively optimizing
for size (in -Oz).  We have seen several uplifts and degradations in performance
of the generated code on several benchmarks, and to avoid the degradations we
decided to only enable hoisting of getElementPtrs without their associated
memory operations when optimizing for size.  This restriction may be lifted in
the future, either when the register allocation rematerialization pass has been
improved to catch the regressions, or by implementing more precise heuristics
for code hoisting.

>>>>

\subsection{Profitability check (Cost models)}
\label{subsec:profitability}
After the legality checks have passed, we check for profitability of hoisting.
That takes into account the impact code-hoisting would have on various
parameters that affect runtime performance e.g., impact on live-range, gain in
the code size.  We have established a set of cost models described in
Section~\ref{sec:cost-models} for each parameter and tuned them for performance
against representative benchmarks.

\begin{verbatim}
    b = m
    c = n
B0: if c is true then goto B3 else goto B4

B1: a0 = b<kill> + c<kill>

B2: a1 = b<kill> + c<kill>

After hoisting a0 and a1 are removed and a copy
of a0 as a01 is placed in B0 just before the branch.
\end{verbatim}

Ideally, it should be okay to hoist all the instructions and a later a
live-range-splitting \cite{cooper1998live} pass should make the right decision
of rematerializing the instruction should it be beneficial to do so. But the
current live-range splitting pass of \LLVM{} is not making the optimial decision
and we have found regressions while hoisting aggressively.

Moreover, LLVM has a `getelementptr' instruction which computes the address
where a load or a store would happen. It is a scalar computation and gets
hoisted frequently even if the loads/stores would not get hoisted.  In order to
reduce register pressure while hoisting loads, we have restricted hoisting of
address computations away from their corresponding loads and stores when the
loads and stores cannot be moved.  This restriction is only to mitigate the
limitations of LLVM's register allocator and may be lifted in the future, when
the register allocation rematerialization pass has been improved to catch the
regressions.

@@

\section{Implementation Details}
\label{sec:implementation-details}

@@
